A Theory for Data Migration of Generated Evolving Information Systems

### 2022-11-20 2022 Migration/articleMigration.tex
verworvenheden van ons gesprek van vandaag.

   A dataset $\dataset$ is a tuple $\pair{\triples}{\inst}$,
   in which $\triples$ is a finite set of triples
   and $\inst$ is the instance relation.
   Where ambiguity lures, we write $\triples_{\dataset}$ and $\inst_{\dataset}$ to indicate that we specifically mean the dataset $\dataset$.
   Every dataset is an element of an infinite set called $\Dataset$.
   To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\triples$.

@@

@@

\subsection{Data migration}
\begin{figure}[bht]
   \begin{center}
     \includegraphics[scale=.45]{Migration.png}
   \end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}
   To migrate system $\infsys$ to $\infsys'$, we take the following steps:
\begin{enumerate}
   \item Let the old system $\infsys$ have state $\dataset$ and schema $\schema$.
         A developer defines a schema $\schema'$ of the new system as a first step in the deployment of the increment.
   \item The generator generates a migration system $\migrsys$ based on $\schema$ and $\schema'$,
         which gives the user the full functionality of the old system and the new system,
         i.e.\ $\schema_{\migrsys}=\schema\sqcup\schema'$.
   \item A migration engineer changes $\migrsys$ to accommodate the non-standard requirements of the migration,
         yielding migration system .
   \item The production environment switches instantly from $\infsys$ to $\migrsys'$,
         while $\dataset_{\infsys}\subseteq\dataset_{\migrsys'}$.
         This means that the data of the old system remains intact.
   \item After a finite span of time, $\dataset_{\migrsys'}$ satisfies all its enforce rules.
         This implies that the generated rules plus the changes made by the migration engineer ensure the preservation of the meaning of the old data.
   \item The system removes the data, the schema, and the functionality of $\migrsys'$ that are not supported by $\schema'$.
         i.e\ $\schema_{\migrsys''}=\schema'$ and $\dataset_{\migrsys''}=\select{\schema'}{\dataset_{\migrsys''}}$.
\end{enumerate}

>>>>

   A dataset $\dataset$ is a tuple $\pair{\triples}{\inst}$,
   in which $\triples$ is a finite set of triples
   and $\inst$ is the instance relation.
   We write $\triples_{\dataset}$ and $\inst_{\dataset}$ to refer to $\triples$ and $\inst$ respectively.
   Every dataset is an element of an infinite set called $\Dataset$.
   To save writing in the sequel, we will write $a\ r\ b$ to denote that $\triple{a}{r}{b}\in\triples$.

@@

   A subset of the rules are so called {\em enforce rules}.
   A developer makes an enforce rule by using some special syntax.
   Whenever the dataset violates an enforce rule,
   the system will make other changes to the dataset to satisfy that rule.
   Hence, an enforce rule $u$ can be violated during a small amount of time in which the system restores its invariance.

@@

\subsection{Data migration}
\begin{figure}[bht]
   \begin{center}
     \includegraphics[scale=.45]{Migration.png}
   \end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}
   To migrate system $\infsys$ to $\infsys'$, we take the following steps:
\begin{enumerate}
   \item Let the old system $\infsys$ have state $\dataset$ and schema $\schema$.
         A developer defines a schema $\schema'$ of the new system as a first step in the deployment of the increment.
   \item The generator generates a migration system $\migrsys$ based on $\schema$ and $\schema'$,
         which gives the user the full functionality of the old system and the new system,
         i.e.\ $\schema_{\migrsys}=(\schema\sqcup\schema')\cup\id{migrationrules}$.
         The schema $\id{migrationrules}$ represents the enforce rules that migrate the existing data to the new dataset.
         TODO: $\id{migrationrules}$ bevat alleen tijdelijke datastructuur, indien nodig.
   \item A migration engineer changes $\id{migrationrules}$ to accommodate the non-standard requirements of the migration,
         yielding migration system $\migrsys'$.
         TODO: de migration engineer kan alleen maar in $\dataset'$ dingen (laten) wijzigen, zodat je altijd terug kunt naar het oude systeem.
   \item The production environment switches instantly from $\infsys$ to $\migrsys'$,
         while $\dataset_{\infsys}\subseteq\dataset_{\migrsys'}$.
         This means that the data of the old system remains intact.
   \item After a finite span of time, $\dataset_{\migrsys'}$ satisfies all its enforce rules.
         This implies that the generated rules plus the changes made by the migration engineer ensure the preservation of the meaning of the old data.
   \item Users may have to perform manual migration steps.
         These are specified as process rules in $\id{migrationrules}$.
         % Dit zijn "kurk" regels.
         The system ensures that the amount of work needed to perform these steps is finite.
   \item The patron of the system can decide whether to carry on with the migration or to roll back to the old system.
         She can do that incrementally by switching off the interfaces one-by-one.
   \item The system retains the data, the schema, and the functionality that is covered by $\schema'$, dropping everything else.
\end{enumerate}
